### 2020-07-31 README.Rmd
:book: Added readme.

---
output:
  github_document:
    html_preview: true
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
```

# Infinite diameter confidence sets in Hedges' publication bias model

Repository for the arXiv version of the paper [Infinite diameter confidence sets in Hedges' publication bias model](https://arxiv.org/abs/1912.09180).

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(out.width='750px', dpi=200)
```