### 2020-08-06 infiniteci.tex
:+1: Small changes to contour plot.

%% LyX 2.3.4.4 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[english]{article}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{graphicx}

\makeatletter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
%% Because html converters don't know tabularnewline
\providecommand{\tabularnewline}{\\}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Textclass specific LaTeX commands.
\numberwithin{equation}{section}
\numberwithin{figure}{section}
\usepackage[natbibapa]{apacite}
\theoremstyle{plain}
\newtheorem{thm}{\protect\theoremname}
\theoremstyle{definition}
\newtheorem{defn}[thm]{\protect\definitionname}
\theoremstyle{definition}
\newtheorem{example}[thm]{\protect\examplename}
\theoremstyle{plain}
\newtheorem{lem}[thm]{\protect\lemmaname}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
\usepackage{infiniteci}

\titletag{An arXiv preprint, v2}
\makeatletter
\titlerunning{\@title}
\makeatother
\authorrunning{Jonas Moss}

\renewcommand{\sqrt}[1]{{(#1)^{1/2}}}
\DeclareMathOperator{\tr}{tr}
\DeclareMathOperator{\Tr}{Tr}
\DeclareMathOperator{\Var}{Var}
\DeclareMathOperator{\argmin}{argmin}
\DeclareMathOperator{\Cor}{Cor}
\DeclareMathOperator{\Cov}{Cov}
\DeclareMathOperator{\diag}{diag} 
\DeclareMathOperator{\argmax}{argmax}
\DeclareMathOperator{\supp}{supp}

\usepackage{enumitem}
\setlist[enumerate]{label = (\roman*)}

\makeatother

\usepackage{babel}
\providecommand{\definitionname}{Definition}
\providecommand{\examplename}{Example}
\providecommand{\lemmaname}{Lemma}
\providecommand{\theoremname}{Theorem}

\begin{document}
\title{Infinite Diameter Confidence Sets in Hedges' Publication Bias Model}
\author{Jonas Moss \orcid{0000-0002-6876-6964}\\
Department of Mathematics, University of Oslo\\
PB 1053, Blindern, NO-0316, Oslo, Norway\emph{}\\
\emph{jonasmgj@math.uio.no}}
\maketitle
\begin{abstract}
Meta-analysis, the statistical analysis of results from separate studies,
is a fundamental building block of science. But the assumptions of
classical meta-analysis models are not satisfied whenever publication
bias is present, which causes inconsistent parameter estimates. Hedges'
selection function model takes publication bias into account, but
estimating and infering with this model is tough. Using a generalized
Gleser--Hwang theorem, we show there is no confidence set of guaranteed
finite diameter for the parameters of Hedges' selection model. This
result provides a partial explanation for why inference with Hedges'
selection model is fraught with difficulties.
\end{abstract}
\keywords{meta-analysis \and confidence intervals \and file-drawer  problem \and publication bias \and selection models \and weigth function models}

\section{Introduction}

A meta-analysis is a statistical analysis that quantiatively combines
results from separate scientific studies. When the studies measure
the same phenomenon, pooling of information allows us to predict the
common effect size with larger precision than we could have done with
one study alone. Meta-analyses are ubiquitous in the empirical sciences
and forms a key component of most systematic reviews such as Cochrane's
reviews \citep{Higgins2019-vv}.

Most meta-analytic techniques assume honest and unbiased reporting
of results. But there is ample evidence that the scientific literature
is not unbiased, as studies with\emph{ }significant \emph{p}-values
tend to be published with a greater probability than other studies
\citep{Easterbrook1991-ph}, a phenomenon called publication bias
by \citet{sterling1959publication} and the file-drawer problem by
\citet{Rosenthal1979-pm}. When publication bias is present, there
is no reason to trust the results of meta-analytic methods that do
not account for it, as the parameter estimates will be incosistent
\citep{Carter2019-rw}.

Hedges' \citeyearpar{hedges1992modeling} publication bias model takes
publication bias explicitly into account using a selection model,
and is arguably the most appropriate model for publication bias \citep{Carter2019-rw}.
Despite there being an $\mathtt{R}$ \citep{Team2013-tt} package
for maximum likelihood estimation of this model, called $\texttt{weightr}$
\citep{Coburn2019-ec}, the model has not yet taken off. Its maximum
likelihood estimation methods are numerically unstable and its estimates
can be off even when they converge \citep{Coburn2019-ec,Stanley2005-ng}.
The estimate of the mean effect size may be negative and of unrealistically
large magnitude, and the estimated heterogeneity parameter might be
improbably large. It turns out there are ridges in the likelihood
function that can be linked to this behavior \citep{McShane2016-rb},
but it has not been stated in clear terms exactly what the consequences
are for inferential procedures. The purpose of this note is to show
there is no confidence set for the mean effect size that has infinite
diameter with probability $0$. In the terminology of \citet{berger1999integrated},
Hedges' publication bias model belongs to the \emph{Gleser--Hwang}
class.

\section{Hedges' publication bias model}

The most popular and well-known meta-analysis method is the random
effects model with normal likelihoods \citep{hedges1998fixed}. Written
in hierarchical notation, it equals
\begin{eqnarray*}
\theta_{i} & \sim & N(\theta_{0},\tau),\\
x_{i}\mid\theta_{i},\sigma & \sim & N(\theta_{i},\sigma_{i}).
\end{eqnarray*}
Here $\sigma_{i}$ is the standard deviation of the $i$th study.
Following the convention in meta-analysis, we assume all $\sigma_{i}$s
to be known. The mean parameter $\theta$ is the population effect
size, $N(\theta_{0},\tau)$ is the effect size distribution, and $\tau$
is the heterogeneity parameter. The purpose of the effect size distribution
is to model the fact that most effect size estimates plugged into
a meta-analysis do not appear to measure the same phenomenon. By integrating
out $\theta_{i}$, we find the density of $x_{i}$,
\[
f(x_{i};\theta_{0},\tau,\sigma_{i})=\phi(x_{i};\theta_{0},\sqrt{\tau^{2}+\sigma_{i}^{2}}),
\]
where $\phi$ is the density of a normal random variable.

We will assume that the random effects meta-analysis model is true
in the absence of publication bias. The mechanisms that cause publication
bias modify the density in a suitable way. Consider the case when
only significant studies at some specified level $\alpha$ are published.
Assuming one-sided tests, the \emph{p}-values are $u_{i}=\Phi(-x_{i}/\sigma_{i})$,
or normal one-sided \emph{p}-values. We will only deal with with one-sided
\emph{p}-values in this paper, as there is usually, but not always,
just one direction that is interesting to researchers, reviewers,
and editors. A one-sided \emph{p}-value can also be used if the researchers
reported a two-sided value, since $p=0.05$ for a two-sided hypothesis
corresponds to $p=0.025$ for a one-sided hypothesis, et cetera.

Define $c_{\alpha}=\Phi^{-1}(1-\alpha)$, the cutoff for significance
at level $\alpha$. The \emph{basic publication bias model} is a truncated
normal model with density
\begin{equation}
f(x_{i};\theta_{0},\sigma_{i})=\Phi\left(\frac{\theta_{0}-c_{\alpha}}{\sqrt{\sigma_{i}^{2}+\tau^{2}}}\right)^{-1}\phi(x;\theta_{0},\sqrt{\sigma_{i}^{2}+\tau^{2}})1[x_{i}/\sigma_{i}>c_{\alpha}],\label{eq:selection for significance model}
\end{equation}
This model for publication bias was introduced by \citet{hedges1984estimation}
in the context of $F$-distributions.

The basic publication bias model is unrealistic. It requires that
no non-significant studies are published. But even in the fields most
severely affected by publication bias, such as psychology, a non-negligble
number of non-significant studies are published \citep{Motyl2017-dx}.
Moreover, the basic publication bias model does not allow for different
cutoffs for significance. For instance, it is likely that some editors
will accept studies reaching a significance at $\alpha=0.025$, corresponding
to $x_{i}/\sigma_{i}>1.96$ but not at $\alpha=0.05$, corresponding
to $x_{i}/\sigma_{i}>1.64$.

These problems can be rectified by adopting the selection model for
publication bias of \citet{iyengar1988selection}, which models the
following scenario.
\begin{quote}
\textbf{Publication bias scenario.} Alice the editor receives a study
with the \emph{p}-value $u$. Her publication decision is a random
function of this \emph{p}-value. That is, she will publish the study
with some probability $w(u)$ and reject it with probability $1-w(u)$.
Every study you will ever read in Alice's journal has survived this
selection mechanism, the rest are lost forever.
\end{quote}
Let $w(u_{i})$ be a function of the \emph{p}-value $u_{i}=1-\Phi(-x_{i}/\sigma_{i})$
taking values in $[0,1]$. Then $w(u_{i})$ is a probability for every
$u_{i}$, and the selection model

\begin{equation}
f(x_{i};\theta_{0},\sqrt{\sigma_{i}^{2}+\tau^{2}})\propto\phi(x_{i};\theta_{0},\sqrt{\tau^{2}+\sigma_{i}^{2}})w(u)\label{eq:iyengar-greenhouse model}
\end{equation}
models the publication bias scenario exactly. This model can be viewed
as a rejection sampling procedure \citep{flury1990acceptance,von1951various},
where $\phi$ serves as proposal distribution for $f$. Variants of
this model, with and without covariates, has been studied by e.g.
\citet{Dear1992-gw,Vevea1995-on,Vevea2005-xp,citkowicz2017parsimonious}.

\citet{hedges1992modeling} studies the selection model when $w$
is a step function with fixed steps. Let $\alpha$ be a vector with
elements $0=\alpha_{0}<\alpha_{1}<\ldots<\alpha_{K}=1$ and $\rho$
be a $K$-ary non-negative, non-increasing vector having its first
element equal to $\rho_{1}=1$. Define the step function $w$ based
on $\alpha$ and $\rho$ as
\begin{equation}
w(u;\rho,\alpha)=\sum_{k=1}^{K}\rho_{k}1_{(\alpha_{k-1},\alpha_{k}]}(u).\label{eq:step function}
\end{equation}
We call the selection model with a step function \emph{Hedges' publication
bias model. }Its density is
\begin{equation}
f(x_{i};\theta_{0},\sqrt{\tau^{2}+\sigma_{i}^{2}})\propto\sum_{k=1}^{K}\rho_{k}\phi(x_{i};\theta_{0},\sqrt{\tau^{2}+\sigma_{i}^{2}})1_{(\alpha_{k-1},\alpha_{k}]}(u_{i}).\label{eq:hedges model}
\end{equation}
Interpretating Hedges' publication bias model is easy. When the editor
receives a study with \emph{p}-value $u$, she finds the $j$ such
that $u\in(\alpha_{j-1},\alpha_{j}]$ and accepts with probability
$\rho_{j}$. Since $\rho_{1}=1$, she always accepts when $u\in[0,\alpha_{1}]$.
The vector $\rho$ is non-increasing since a publication decision
based solely on \emph{p}-values should always act favorably towards
lower \emph{p}-values. The parameters $(\theta_{0},\tau,\rho)$ of
the model are identified when $\alpha$ is fixed \citep[Web Appendix A]{moss2019modelling}.

Hedges' publication bias models allows both for non-significant studies
to be published and allows the editor to act differently towards different
cutoffs such as $\alpha=0.025$ and $\alpha=0.05$. In addition, the
model can approximate any non-increasing selection function $w$ by
increasing the number of steps.

We can write Hedges' model as a mixture model on the form
\begin{equation}
f(x_{i};\theta_{0},\sqrt{\tau^{2}+\sigma_{i}^{2}})=\sum_{k=1}^{K}\pi^{k}f^{k}(x_{i};\theta_{0},\sqrt{\tau^{2}+\sigma_{i}^{2}}).\label{eq:mixture model formulation}
\end{equation}
Here $f^{K}$ is a normal density, $f^{k},k<K$ are normal densities
truncated to $[\Phi^{-1}(1-\alpha_{j-1}),\Phi^{-1}(1-\alpha))$, and
$\pi^{k}$ are mixture probabilities, i.e., $\pi^{k}>0$ for each
$k$ and $\sum_{k=1}^{K}\pi^{k}=1$. The mixture probabilities $\pi^{k}$
are functions of $(\theta_{0},\tau,\sigma_{i},\rho)$, see the appendix
(p. \pageref{eq:pi_i formula}) for their formula.

The main benefit of Hedges' publication bias model (\ref{eq:hedges model})
is how it models \emph{p}-values based publication bias directly,
there is no approximation involved. If you believe in the random effects
meta-analysis model and the \emph{p}-value based publication bias
scenario, Hedges' publication bias model is simply the correct model.
Most statistical methods correcting for publication bias in the literature
either do not make use of an explicit statistical model or do not
estimate the parameters $\theta_{0}$ and $\tau$. For instance, the
funnel funnel plot of \citet{Egger1997-rd} is a graphical method,
while the trim-and-fill method of \citet{Duval2000-ct} is a non-parametric
method based on making the funnel plot symmetric. \citet{Stanley2005-ng,stanley2014meta}
discuss various misspecified regression-based estimators of the corrected
effect size $\theta$ based on the fixed effect Hedges' publication
bias model. The estimating \emph{p}-curve method of \citet{Simonsohn2014-cn}
and \emph{p}-uniform of \citet{Van_Assen2015-qs,Van_Aert2016-cu}
are two methods for dealing with publication bias hailing from psychology.
Both are based on a variant of the basic publication bias model, but
with fixed instead of random effects, and both employ somewhat unusual
estimation methods \citep{McShane2016-rb}. Since there is ample evidence
of heterogeneity in meta-analysis, restricting oneself to the fixed
effects meta-analysis is a mistake.

Hedges' model has some downsides. It models only bias due to selection
of \emph{p}-values, not every source of bias, such as language bias
\citep{Egger1998-kj}. Second, it may not be best model for biases
with other causes than the publication process, such as \emph{p}-hacking
\citep{simmons2011false}. \citet{moss2019modelling} propose a related
model that may be more successful at correcting for \emph{p}-hacking.

Estimation of Hedges' model is difficult. \citet[Section 6.3]{Stanley2005-ng}
discusses three problematic cases in economics when maximum likelihood
was used to estimate Hedges' publication bias model. \citet[Appendix A]{McShane2016-rb}
notes that while estimation of the basic publication bias model is
hard, introducing the heterogeneity parameter exacerbates the problem.
The likelihood function has contours following approximately $\tau\propto\sqrt{|\theta|}$.
Figure \ref{fig:contour lines} shows the contour lines for the meta-analysis
of \citet{Cuddy2018-kp} where the selection probabilities of the
step function (\ref{eq:step function}) are fixed at $\rho=(1,0.6,0.1)$.
Only just around the maximum at $\hat{\theta}_{0}=0.55$ and $\hat{\sigma}=5\cdot10^{-7}$
can the likelihood be approximated with a quadratic function.

\begin{figure}
\noindent \begin{centering}
\includegraphics[scale=0.5]{chunks/cuddy}
\par\end{centering}
\caption{\label{fig:contour lines}Contour lines for the log-likelihood for
the simple publication model using power posing data of \citet{Cuddy2018-kp}. }

\end{figure}

\section{Confidence sets of infinite diameter}

Fix some measurable space $(\Omega,\mathcal{F})$, let $\mathcal{P}$
be a family of dominated probability measures defined on this measurable
space, and $\mathcal{\Pi}$ a partition of $\mathcal{P}$. Recall
that a partition of $\mathcal{P}$ is a collection of disjoint non-empty
subsets $\pi$ of $\mathcal{P}$ such that $\bigcup_{\pi\in\Pi}\pi=\mathcal{P}$.
When $p$ is a density associated with a $P\in\mathcal{P}$, we will
use the standard notation $[p]$ to denote the unique $\pi$ containing
$P$. 
\begin{defn}
\label{def:confidence set} A \emph{confidence set} of level $\alpha$
is a family of rejection sets $\{R(\pi)\},\pi\in\Pi$ such that
\[
\sup_{\pi\in\Pi}\sup_{P\in\pi}P(R(\pi))\leq\alpha.
\]
If the inequality is an equality, the confidence set has \emph{size}
$\alpha$. 
\end{defn}

This definition of confidence sets might look slightly unfamiliar,
but it is a straight-forward generalization of the definitions in
\citet[Definition 9.1.5]{Casella2002-lg} and \citet[Section 3.5]{lehmann2006testing}.
Usually, a confidence is defined as a set $C$ adhering to the relation
\begin{equation}
\pi_{0}\in C\iff\omega\notin R(\pi_{0}).\label{eq:c-confidence set}
\end{equation}
That is, $\pi_{0}\in C$ if and only if we accept the null-hypothesis
$H_{0}:\pi=\pi_{0}$, but will only need the formulation using rejection
sets in this paper. When confidence sets are defined in terms of rejection
sets, there is sometimes no partition $\Pi$ to take the supremum
over, and the definition reduces to $\sup_{P\in\mathcal{P}}P(R(P))\leq\alpha$.
The term\emph{ confidence interval} is far more common than confidence
set, but this requires that $C$ of equation \ref{eq:c-confidence set}
is an interval, which we will not require here.

The following example should make Definition \ref{def:confidence set}
familiar.
\begin{example}
\label{exa:t-test} Consider the usual \emph{t}-confidence interval
with $n$ observations. Here $\mathcal{P}$ contains all measures
$P_{\mu,\sigma}^{n}$, where $P_{\mu,\sigma}$ is the probability
measure of a normal with mean $\mu$ and standard deviation $\sigma$.
Furthermore, when $Q$ is a probability measure, $Q^{n}$ denotes
the $n$-fold product measure of $Q$, corresponding to $n$ independent
samples from $Q$. Let $\pi(\mu)=\{P_{\mu,\sigma}^{n}\mid\sigma>0\}$
contain all normal probability measures with mean $\mu$ and some
positive standard deviation. Then $\{\pi(\mu)\},\mu\in\mathbb{R}$
defines a partition of $\mathcal{P}$. A two-sided \emph{t}-confidence
interval is a confidence set of size $\alpha$ with partition $\{\pi(\mu)\},\mu\in\mathbb{R}$
according to definition \ref{def:confidence set}.
\end{example}

Now we must find out how to measure the size of confidence sets. Let
a \emph{size function} be any positive function $\left\Vert \cdot\right\Vert :\Pi\to[0,\infty)$.
In the \emph{t}-confidence interval example above, $\left\Vert \cdot\right\Vert $
can be taken to be $\left\Vert \pi\right\Vert =|\mu|$ for the unique
$\mu$ associated with each $\pi$.

The diameter of a confidence set is the random variable
\begin{equation}
D(\omega)=\sup_{\pi\in\Pi}\{\left\Vert \pi\right\Vert \mid\omega\notin R(\pi)\}.\label{eq:diameter}
\end{equation}
That is, the diameter tells you the size of the largest accepted $\pi$.
We will assume that $D$ is Borel measurable.
\begin{defn}
\label{def:infinite diameter}A confidence set has infinite diameter
with $P$-positive probability if $P(D=\infty)>0$. It has infinite
diameter with positive probability if $P(D=\infty)>0$ for all $P\in\mathcal{P}$.
\end{defn}

This is a modified Gleser--Hwang theorem \citep[Theorem 1]{gleser1987nonexistence}
formulated in terms of partitions. Its proof is in the appendix (p.
\pageref{proof:Gleser--Hwang}).
\begin{thm}[Gleser--Hwang theorem]
\label{thm:Gleser--Hwang} Suppose there is a sequence $\{p_{i}\}$
of densities from $\mathcal{P}$ satisfying the following: 
\begin{enumerate}
\item There is a density $p^{\star}$ such that $p_{i}$ converges to $p^{\star}$
pointwise,
\item $\supp p\supseteq\supp p^{\star}$ for all $p$,
\item the size of $[p_{i}]$ goes to infinity, $\left\Vert [p_{i}]\right\Vert \to\infty$.
\end{enumerate}
Then every confidence set with level $\alpha>0$ has infinite diameter
with positive probability.
\end{thm}

Following the terminology of \citet{berger1999integrated}, we will
say that families $\mathcal{P}$ of probabilities satisfying the conclusion
of Theorem \ref{thm:Gleser--Hwang} for some suitable partition $\Pi$
belong to the Gleser--Hwang class. To make the Gleser--Hwang class
more familiar, we will present two examples. More examples can be
found in the papers of \citet{gleser1987nonexistence} and \citet{berger1999integrated}.
\begin{example}[Fieller's problem]
\label{exa:Fieller's problem} Let $X,Y$ be an observation from
a bivariate normal $N([\mu_{1},\mu_{2}],I\sigma^{2})$, where $\sigma^{2}$
is known. We want to form a confidence set for ratio $\mu_{2}/\mu_{1}$.
The most famous confidence set is due to \citet{Fieller1940-lg}.
His confidence set can be finite, the whole real line, or the union
of two disjoint semi-infinite intervals \citep{Koschat1987-dk}, all
with positive probability.

But it is not only Fieller's confidence set that might be infinitely
long. The Gleser--Hwang theorem can be used to show that every confidence
set for $\mu_{2}/\mu_{1}$ must be infinitily long with positive probability.
This result is almost independent of the distribution of $X$ and
$Y$. To state this result in our notation, let $\mathcal{P}$ be
a family of bivariate distributions over $(X,Y)$. All of these distributions
have the same support, and all of them have finite means $E_{p}(X)$
and $E_{p}(Y)$. Moreover, assume $E_{p}(X)=0$ is attainable for
some $p\in\mathcal{P}$. Define the partition $\Pi$ by $p,q\in\pi$
if and only if $E_{p}(X)/E_{p}(Y)=E_{q}(X)/E_{q}(Y)$, and let $\left\Vert [p]\right\Vert =|E_{q}(X)/E_{q}(Y)|$,
i.e., the ratio of means. Choose a sequence $p_{n}(x,y)=p(x,y)$,
where $p(x,y)$ is density with means $E_{p}X>0$ and $E_{p}Y=0$.
Then $\left\Vert [p]\right\Vert =\infty$, the conditions of Theorem
\ref{thm:Gleser--Hwang} are satisfied, and every confidence set with
level $\alpha>0$ has infinite diameter with positive probability.
\begin{example}[Bahadur--Savage]
\label{exa:Bahadur--Savage} \citet{bahadur1956nonexistence} studies
non-parametric testing of the mean, and concludes the mean cannot
be meaningfully tested. They are working with a family $\mathcal{P}$
of densities over $\mathbb{R}$ that covers all finite means, has
finite variances, and is closed under convex combinations. Similar
problems were considered by \citet{romano2004non} and \citet{Donoho1988-hg}.

Using the Gleswer--Hwang theorem, it is easy to verify every confidence
set has infinite diameter with positive probability. Define the partition
$\Pi$ by $p,q\in\Pi$ if and only if $E_{p}(X)=E_{q}(X)$, and let
$\left\Vert [p]\right\Vert =|E_{p}(X)|$. Let
\[
p_{n}(x)=\left(1-\frac{1}{n}\right)q_{0}(x)+\frac{1}{n}q_{n^{2}}(x),
\]
where $p_{0}$ has mean $0$ and $q_{n^{2}}$ has mean $n^{2}$. Then
$\left\Vert [p_{n}]\right\Vert =n$, the conditions of Theorem \ref{thm:Gleser--Hwang}
are satisfied, and every confidence set with level $\alpha>0$ has
infinite diameter with positive probability.
\end{example}

\end{example}

There are several natural candidates for $\Pi$ when working with
Hedges' selection function model (\ref{eq:hedges model}). We will
work with three of them. First, consider the partition where $p,q\in\pi$
if and only if $p,q$ have the same mean effect size parameter $\theta_{0}$.
We will equip this partition with the size function $\left\Vert \pi\right\Vert =|\theta_{0}|$,
and it corresponds to a confidence set for $\theta_{0}$. Second,
consider the partition where all $p,q\in\pi$ have the same heterogeneity
parameter $\tau$, equipped with $\left\Vert \pi\right\Vert =\tau$.
Finally, we will work with the partition where all $p,q\in\pi$ have
the same heterogeneity parameter $\tau$ and population effect size
$\theta_{0}$, and equip it with $\left\Vert \pi\right\Vert =\sqrt{\theta_{0}^{2}+\tau^{2}}$.
This information is summarised in Table \ref{tab:Possible partitions}
for convenience.

\begin{table}
\noindent \centering{}\caption{\label{tab:Possible partitions}The three partitions $\Pi$ for the
selection function model}
{\small{}}%
\begin{tabular}{llll}
 & {\small{}Symbol} & {\small{}Size $\left\Vert \cdot\right\Vert $} & {\small{}Confidence set}\tabularnewline
\hline 
{\small{}Mean effect size} & {\small{}$\theta_{0}$} & {\small{}$\left\Vert \pi\right\Vert =|\theta_{0}|$} & {\small{}Confidence set for $\theta_{0}$}\tabularnewline
{\small{}Heterogeneity parameter} & {\small{}$\tau$} & {\small{}$\left\Vert \pi\right\Vert =\tau$} & {\small{}Confidence set for $\tau$}\tabularnewline
Both parameters & {\small{}$(\theta_{0},\tau)$} & {\small{}$\left\Vert \pi\right\Vert =\sqrt{\theta_{0}^{2}+\tau^{2}}$} & {\small{}Joint confidence set for $(\theta,\tau)$}\tabularnewline
\end{tabular}
\end{table}

Let us take a look at the basic publication bias model (\ref{eq:selection for significance model})
again. To use Theorem \ref{thm:Gleser--Hwang} we need a witnessing
sequence of functions $p_{n}\to p$ satisfying the conditions $\text{(ii)}$
and $\text{(iii)}$. The next lemma shows how to make such a witness
for the truncated normal. Its proof is in the appendix (p. \pageref{proof:generalized shifted exponential}).
\begin{lem}
\label{lem:generalized shifted exponential} Let $f_{n}$ be a normal
density truncated to $[a,b)$, where $b=\infty$ is allowed, with
underlying mean $\theta_{n}=-n$ and standard deviation $\sigma_{n}^{2}=n^{2}+c$
for some $c\in\mathbb{R}$. Then $f_{n}$ converges pointwise to $\exp(-x)/[\exp(-a)-\exp(-b)]$,
the distribution of an exponential variable truncated to $[a,b)$.
\end{lem}

Using Lemma \pageref{lem:generalized shifted exponential} it is not
hard to show that the basic publication bias model (\ref{eq:selection for significance model})
is a member of the Gleser--Hwang class.
\begin{thm}
\label{thm:simple publication bias} Assume $n$ independent samples
from the basic publication bias model (\ref{eq:selection for significance model}).
Then, for any $n$, any confidence set for $\theta_{0},\tau$, or
$(\theta_{0},\tau)$ with level $\alpha>0$ will have infinite diameter
with positive probability. 
\end{thm}

\begin{proof}
\label{proof:simple publication bias}Let $\Pi$ be the partition
of $\mathcal{P}$ where $p,q\in\pi$ if and only if they share the
same $\theta_{0}$, and let $||[p]||=|\theta_{0}|$. We are dealing
with products of densities of the form (\ref{eq:selection for significance model}),
that is,
\[
p(x)=\prod_{i=1}^{N}\Phi\left(\frac{\theta_{0}-c_{\alpha}}{\sqrt{\sigma_{i}+\tau}}\right)^{-1}\phi(x_{i};\theta_{0},\sqrt{\sigma_{i}^{2}+\tau^{2}}),
\]
where $\sigma_{i}$ are known parameters. From Lemma \ref{lem:generalized shifted exponential},
$p_{n}$ converges to a product of truncated exponentials when $\theta_{n}=-n$
and $\tau_{n}^{2}=n$. Since $||[p_{n}]||=n$, the three conditions
of Theorem \ref{thm:Gleser--Hwang} are satisfied. The proofs for
$||[p]||=\tau$ and $||[p]||=\sqrt{\tau^{2}+\theta_{0}^{2}}$ are
similar and omitted.
\end{proof}
Proving the analogue of Theorem \ref{thm:simple publication bias}
for Hedges' publication is only somewhat more involved. We will use
the mixture representation of (\ref{eq:mixture model formulation})
and a lemma generalizing Theorem \ref{thm:Gleser--Hwang} to a certain
kind of mixtures.

Let $f^{1},f^{2},\ldots,f^{K}$ be a sequence of densities, $\pi=(\pi_{1},\pi_{2},\ldots,\pi_{K})$
a probability vector, and $p=\sum_{k\leq K}\pi^{k}f^{k}$ a mixture
distribution. We will assume that the size of $[p]$ equals the size
of any of its mixture components $[f^{k}]$ for some size $\left\Vert \cdot\right\Vert $,
i.e., $\left\Vert [p]\right\Vert =\left\Vert [f^{k}]\right\Vert $
for all $k$. Why we do this will be clear in the proof of Theorem
\ref{thm:general publication bias}, but think of it this way: If
all of $p$s mixture components have the same mean, the mean of $p$
equals the mean of any $f^{k}$. 
\begin{lem}
\label{prop:Mixture model corollary} Let $\mathcal{P}$ be a class
of $K$-ary mixture distributions and $||[p]||$ be the assumption
above. Assume there is a sequence $p_{i}=\sum_{k\leq K}\pi_{i}^{k}f_{i}^{k}$
and a subset $K'$ such that 
\begin{enumerate}
\item For all $k\in K'$, there is a density $f^{k\star}$ such that $f_{i}^{k}$
converges to $f^{k\star}$ pointwise.
\item For all mixtures $p$, $\supp p\supseteq\supp f^{k\star}$ for some
$k\in K'$.
\item For all $k\in K'$, the size of $[f_{i}^{k}]$ goes to infinity, $\left\Vert [f_{i}^{k}]\right\Vert \to\infty$.
\item The density concentrates on the components indexed by $K'$, $\lim_{i\to\infty}\sum_{k\in K'}\pi_{i}^{k}=1$.
\end{enumerate}
Then every confidence set with level $\alpha>0$ has infinite diameter
with positive probability.
\end{lem}

\begin{proof}
We employ Theorem \ref{thm:Gleser--Hwang}. By (i) and (iv), $p_{i}$
converges pointwise to the density $\sum_{k\in K'}\pi_{i}^{k\star}f_{i}^{k\star}=p^{\star}$
. That $\supp p\supseteq p^{\star}$ follows from (ii) and (iv). Finally,
from the assumption that$\left\Vert [p]\right\Vert =\left\Vert [f_{k}]\right\Vert $,
we get that $\left\Vert [p]\right\Vert \to\infty$ too. 
\end{proof}
\begin{thm}
\label{thm:general publication bias} Assume $n$ independent samples
from the publication bias model (\ref{eq:selection for significance model}),
where the selection probability $\rho$ is unknown and $\alpha$ is
known. Then any confidence set for $\theta_{0},\tau$, or $(\theta_{0},\tau)$
will have infinite diameter with positive probability.
\end{thm}

\begin{proof}
\label{proof:general publication bias} Let $n=1$ and consider confidence
sets for $\theta_{0}$. Let $\Pi$ be the partition of $\mathcal{P}$
where $p,q\in\pi$ if and only if they share the same $\theta_{0}$,
and let $||[p]||=|\theta_{0}|$. Then $||f^{k}||=|\theta_{0}|$ from
the mixture representation (\ref{eq:mixture model formulation}).
Using Lemma \ref{lem:generalized shifted exponential}, we see that
$f^{k},k>1$ converges pointwise to a truncated exponential when $\theta_{0}=-n$
and $\sqrt{\tau^{2}+\sigma^{2}}=n^{1/2}$, so that (i), (ii), (iii)
of Proposition \ref{prop:Mixture model corollary} are satisfied with
$K'=K\backslash\{1\}$. The mixture probabilities for $k\neq K$ can
be fixed at e.g. $\pi=1/(K-1)$, and (iv) is satisfied as well. The
proofs for $||[p]||=\tau$ and $||[p]||=\sqrt{\tau^{2}+\theta_{0}^{2}}$
are similar and omitted. When $n>1$, expand the expression $\prod_{j=1}^{n}\sum_{k\leq K}\pi_{i}^{k}(\sigma_{j})f_{i}^{k}(\sigma_{j}),$and
use the same reasoning as in the first part of this proof.
\end{proof}

\section{Remarks}

Well-behaved confidence sets for Hedges publication bias model do
not exist, but well-behaved credibility sets do. Bayesian estimation
of Hedges' model can be made routine, as it is easy to find uncontroverisal
priors for $\theta_{0}$ and $\tau$. In practical meta-analyses we
know that $\theta_{0}$ cannot be large, and is likely to be close
to $0$. Moreover, since it is common effects in meta-analyses to
be intepreted as the aggregation of many small effects, the central
limit theorem justifies using a normal prior. As we want to remove
prior mass from negative $\theta_{0}$s of large magnitude, $N(0,1)$
is a decent standard prior. Similarily, a half-normal is a reasonable
prior for the heterogeneity parameter $\tau$. \citet{moss2019modelling}
employed these priors on several examples.

\section*{Appendix}

The following sandwich convergence theorem is used in the proof of
the Gleser--Hwang theorem. 
\begin{lem}[{\citet[Exercise 16.4(a)]{billingsley1995probability}}]
\label{lem:Dominated covergence theorem} Suppose $a_{n},b_{n},f_{n}$
converge to $a,b,f$ and $a_{n}\leq f_{n}\leq b_{n}$ for all $n$.
If $\int a_{n}d\mu\to\int ad\mu$ and $\int b_{n}d\mu\to\int bd\mu$,
then $\int f_{n}d\mu\to\int fd\mu$.
\end{lem}

The proof of Theorem \ref{thm:Gleser--Hwang} follows the proof of
\citet[Theorem 1]{gleser1987nonexistence} closely. 
\begin{proof}[Proof of Theorem \ref{thm:Gleser--Hwang}]
\label{proof:Gleser--Hwang}We can assume without loss of generality
that $\left\Vert [p_{n}]\right\Vert \geq n$, as we can choose a subsequence
if we have to. By definition of the diameter $D$ (\ref{eq:diameter})
we see that
\[
\{D\geq n\}=\{\omega\in\Omega\mid\textrm{there is a }\pi\textrm{ such that}\left\Vert \pi\right\Vert \geq n\textrm{ and }\omega\in R^{c}(\pi)\},
\]
and if $||[p_{n}]||\geq n$, then $R^{c}([p_{n}])\subseteq\{D\geq n\}$.
Since we assume that $\left\Vert [p_{n}]\right\Vert \geq n$ and 
\[
1-\alpha\leq P_{n}(R^{c}([p_{n}]))=\int_{R^{c}([p_{n}])}p_{n}d\mu
\]
by definition of a confidence set, we have that
\begin{equation}
0<1-\alpha\leq\int_{R^{c}([p_{n}])}p_{n}d\mu\leq\int_{D\geq n}p_{n}d\mu\label{eq:inequality with R^c}
\end{equation}
for all $n$. Since $p_{n}$ and $p^{\star}$ are densities,
\[
\lim_{n\to\infty}\int p_{n}d\mu=1=\int p^{\star}d\mu=\int\lim_{n\to\infty}p_{n}d\mu.
\]
This allows us to use Lemma \ref{lem:Dominated covergence theorem}
with $a_{n}=0$, $b_{n}=p_{n}$, and $f_{n}=1_{D\geq n}p_{n}$ and
conclude that

\begin{equation}
\int_{D\geq n}p_{n}d\mu\to\int_{D=\infty}p^{\star}d\mu.\label{eq:limit of pn}
\end{equation}
Combining equations (\ref{eq:inequality with R^c}) and (\ref{eq:limit of pn}),
we get
\[
0<1-\alpha\leq\int_{D=\infty}p^{\star}d\mu.
\]
Let $P\in\mathcal{P}$ be arbitrary, $p$ be its density, and consider
\[
P(D=\infty)=\int_{D=\infty}pd\mu\geq\int_{D=\infty\cap\supp p^{\star}}\left(\frac{p}{p^{\star}}\right)p^{\star}d\mu.
\]
Since $\int_{D=\infty}p^{\star}d\mu>0$ and $p/p^{\star}>0$ on $\supp p^{\star}$
(since $\supp p\supseteq\supp p^{\star}$ by assumption), we see that
$\int_{D=\infty\cap\supp p^{\star}}(p/p^{\star})p^{\star}d\mu>0$
too. It follows that $P(D=\infty)>0$, and, since $P$ is aribtrary,
$D$ has infinite diameter with positive probability.
\end{proof}
Now we prove Lemma (\ref{lem:generalized shifted exponential}).
\begin{proof}[Proof of Lemma (\ref{lem:generalized shifted exponential})]
\label{proof:generalized shifted exponential}Let $n^{2}>-c$, so
that $\sigma_{n}^{2}>0$. Recll the well-known formula for the normal
truncated to $[a,b]$,
\begin{eqnarray}
f_{n}(x) & = & \frac{1}{\Phi\left(\frac{b-\theta_{n}}{\sigma_{n}}\right)-\Phi\left(\frac{a-\theta_{n}}{\sigma_{n}}\right)}\phi(x;\theta_{n},\sigma_{n})1[a,b](x),\label{eq:truncated}\\
 & = & \frac{\phi(x;-n,(n+c)^{1/2})1[a,b](x)}{\Phi[-(a+n)(n+c)^{-1/2}]-\Phi[-(b+n)(n+c)^{-1/2}]}.\nonumber 
\end{eqnarray}
The normal density part equals
\[
\phi(x;-n,(n+c)^{1/2})=(2\pi)^{-1/2}(n+c)^{-1/2}\exp\left(-\frac{x^{2}+2nx+n^{2}}{2(n+c)}\right).
\]
When $n$ is large compared to $x$, the term $x^{2}/2(n+c)$ is negligible,
hence
\begin{eqnarray*}
\phi(x;-n,(n+c)^{1/2}) & \approx & (2\pi)^{-1/2}(n+c)^{-1/2}\exp(-n^{2}/2(n+c))\exp(-x),\\
 & = & (n+c)^{-1/2}\phi(n/(n+c)^{1/2})\exp(-x).
\end{eqnarray*}
From Equation 5 of \citet{borjesson1979simple}, we know that $\Phi(-x)\approx\phi(x)/x$
as $x$ grows. The part of (\ref{eq:truncated}) involving cumulative
normal distributions become
\begin{eqnarray*}
\Phi[-(a+n)(n+c)^{-1/2}] & \approx & \frac{(n+c)^{1/2}}{n+a}\phi[-(a+n)(n+c)^{-1/2}],
\end{eqnarray*}
and using the same reasoning as above, we find that $\phi[-(a+n)(n+c)^{-1/2}]\approx\phi((n+c)^{1/2})\exp(-a)$
as $n$ increase. Therefore,
\[
\Phi[-(a+n)(n+c)^{-1/2}]\to\frac{(n+c)^{1/2}\phi((n+c)^{1/2})\exp(-a)}{a+n}.
\]
Since this reasoning applies to $b$ as well, we get that $f$ approaches
\begin{eqnarray*}
 &  & \frac{(n+c)^{-1/2}\phi(n/(n+c)^{1/2})\exp(-x)}{\Phi[-(a+n)n^{-1/2}]-\Phi[-(b+n)n^{-1/2}]},\\
 & \approx & \frac{(n+c)^{-1/2}\phi(n/(n+c)^{1/2})\exp(-x)}{(n+c)^{1/2}\phi((n+c)^{1/2})\left[\frac{\exp(-a)}{a+n}-\frac{\exp(-b)}{b+n}\right]},\\
 & = & \frac{\phi(-n^{2}/2(n+c))\exp(-x)}{\phi((n+c)^{1/2})\left[\frac{\exp(-a)}{a+n}-\frac{\exp(-b)}{b+n}\right]},\\
 & \approx & \exp(-x)n^{-1}\left[\frac{\exp(-a)}{a+n}-\frac{\exp(-b)}{b+n}\right]^{-1},\\
 & \to & \exp(-x)\left[\exp(-a)-\exp(-b)\right]^{-1}.
\end{eqnarray*}
Here the third line follows from $\phi(n/(n+c)^{1/2})/\phi((n+c)^{1/2})\to1$.
\end{proof}
For convenience, we show the formula for the mixture probabilities
$\pi_{i}$ here. Let $c_{j}=\Phi^{-1}(1-\alpha_{j})$ and define 
\[
c=\sum_{k=1}^{K}\rho_{k}[\Phi(c_{j-1};\theta_{0},(\tau^{2}+\sigma_{i}^{2})^{1/2})-\Phi(c_{j};\theta_{0},(\tau^{2}+\sigma_{i}^{2})^{1/2})].
\]
Then $\pi_{K}=\rho_{K}c^{-1}$ and \label{eq:pi_i formula}
\[
\pi_{k}=c^{-1}(\rho_{k}-\rho_{K})[\Phi(c_{k-1};\theta_{0},(\tau^{2}+\sigma_{i}^{2})^{1/2})-\Phi(c_{k};\theta_{0},(\tau^{2}+\sigma_{i}^{2})^{1/2})]
\]

\bibliographystyle{apacite}
\bibliography{infiniteci}

\end{document}